Exact asymptotic expressions for local clustering in hyperbolic random graphs \\ \TM{ I prefer not to use the term ``hyperbolic random graph". Better : KPKVB random graph or, in the title,``a hyperbolic model of complex networks"}

### 2019-01-10 concentration_argument.tex
Merging Tobias' comments

Many of the computations in this paper will involve integrals whose integrand contains $\Prob{\Po(\mu_n(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and $\mu_n(y) \approx e^{y/2}$. For instance, the expected modified local clustering function $c_\infty(k)$ in $G_\Pcal$ equals
\[
	\frac{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \Delta_{\Pcal}(y) \alpha e^{-\alpha y} \dd y}
	{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \alpha e^{-\alpha y} \dd y},
\]
where $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$.

@@

\subsection{Concentration argument for the infinite model}

Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xi_{\alpha,\nu}e^{\frac{y}{2}}$. 
For any sequence $\{k_n\}_{n \ge 1}$, possibly constant, define
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\end{equation}
and define further, for any $C > 0$
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(k_n) = \left\{p=(x,y) \in \R \times \R_+ : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
	\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},
\end{equation}
Where we recall that $a \vee b = \max\{a,b\}$ and $a \wedge b = \min\{a,b\}$.

Note that if $k_n = \Omega(\log(n))$ then $e^{y/2} = \bigT{k_n}$ whenever $p=(x,y) \in \Kcal_C(k_n)$. The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{\Rcal_n} \rho(y,k_n) h(y) f_{\alpha,\nu}(x,y) \dd x \dd y
\]
it is enough to consider integration of $\Kcal_C(k_n)$ instead of $\Rcal_n$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \rightarrow  \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = \ell_n \pm C \kappa_n, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right),
\] 
we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd x \dd y
	&= \begin{cases}
		\bigO{\log(n)^{k_n} n^{-\frac{C}{2}}} &\mbox{if } k_n = \bigT{1} \\
		\bigO{k_n^{-(1+C^2)/2}} &\mbox{else.}
	\end{cases}
\end{aligned}
\end{equation}
as $n \to \infty$.

@@

\begin{proof}
Note that in this case
\[
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{\ell_n \log(\ell_n)} &\mbox{else.}
	\end{cases}
\]
Then, since $a_n^- < k_n$ and $a_n^+ > k_n$, $\rho_{y}(k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho_{k_n}(y) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho_{k_n}(y) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{R_n} e^{\beta y} \rho_{k_n}(y) \alpha e^{-\alpha y}x \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho_{k_n}(y) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{R_n} \rho_{k_n}(y) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(y,a_n^+) \int_{a_n^+}^{R_n} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Note that if $k_n = \bigT{1}$ then, for large enough $n$, the first integral is zero. We conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho_{k_n}(y) \alpha e^{-\alpha y} \dd y
	&= \begin{cases}
		\bigO{1} \rho(a_n^+,k_n) &\mbox{if } k_n = \bigT{1},\\
		\bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) &\mbox{else.}
	\end{cases}	
\end{aligned}
\end{equation}

We shall now bound the terms $\rho(a_n^\pm,k_n)$, starting with $\rho(a_n^+,k_n)$.  Using that $k! > \sqrt{2\pi} k^{k + 1/2} e^{-k}$ we write
\begin{align*}
	\rho(a_n^+,k_n) &\le \frac{\Mu{\BallPo{a_n^+}}^{k_n}}{k_n!} e^{-\Mu{\BallPo{a_n^+}}} \\
	&\le (2\pi)^{-1/2} k_n^{-1/2} \left(\frac{\Mu{\BallPo{a_n^+}}}{k_n}\right)^{k_n} e^{-(\Mu{\BallPo{a_n^+}} - k_n)}\\
	&= (2\pi)^{-1/2} k_n^{-1/2} 
		e^{-k_n\left(\frac{\Mu{\BallPo{a_n^+}}}{k_n} - 1 - \log\left(\frac{\Mu{\BallPo{a_n^+}}}{k_n}\right)\right)}.
\end{align*}

Let us first consider the case $k_n \to \infty$, in which case $\kappa_n = \sqrt{\ell_n \log(\ell_n)}$. Since 
\[
	\frac{\Mu{\BallPo{a_n^+}}}{k_n} = \frac{\lambda_n^-}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
Let use now consider the case $k_n = \bigT{1}$. Then $\kappa_n = \log(n)$,
\[
	\frac{\Mu{\BallPo{a_n^+}}}{k_n} = \frac{\lambda_n^+}{k_n} = 1 + \epsilon_n + C \frac{\log(n)}{k_n},
\] 
and hence
\begin{equation}\label{eq:concentration_lemma_bound_an+_constant}
	\rho(a_n^-,k_n) \le \bigO{\log(n)^{k_n} n^{-C}}.
\end{equation}

@@

\begin{remark}[Concentration argument]\label{rmk:concentration_argument}
Lemma~\ref{lem:concentration_argument} will prove very useful in the remainder of this paper since we often have to deal with integrands of the form $g_n(y) f_{\alpha,\nu}(x,y)$ where $g_n(y) = \bigO{n^{-1} k_n^{s}}e^{\beta y}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$. In this case the lemma tells us that for a suitable $C > 0$ we only need to integrate over $\Kcal_C(k_n)$. In order words, we may always assume for $g_n(y)$ (for a penalty of $\smallO{1}$) that $k_n - C \kappa_n \le \xi_{\alpha,\nu} e^{y/2} \le k_n + C \kappa_n$. We will refer to this as a \emph{concentration argument}, e.g. \emph{by a concentration argument}
\[
	\int_{\Rcal_n} g_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y 
	= (1+\smallO{1}) \int_{\Kcal_C(k_n)} g_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y.
\]
\PvdH{ @All:It might be a good idea to add one more example of how the concentration argument is used in the paper. I someone finds a good example we use later on, please add it here.}
\end{remark}

@@

Recall that under the coupling between the hyperbolic random graph and the finite box model $|x-x^\prime|_{\pi e^{R_n/2}} \le \Omega(r,r^\prime)$, while the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Omega(r, r^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for $y + y^\prime < R_n$. This result enables us to determine the measure of a ball around a given point $p=(0,y)$ which will be fairly useful in our subsequent analysis.

\begin{lemma}\label{lem:average_degree_hyperbolic}
Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
\[
	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
\]
where
\[
	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 1 - 4\pi}{4\pi} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
		+ \left(\alpha - \frac{1}{2}\right)\pi e^{-(\alpha - \frac{1}{2})R_n - y/2},
\]
and
\[
	\phi_{\H,n}^{(2)}(y) = 
	+ \begin{cases}
			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
			&\mbox{if } 1/2 < \alpha < 3/2,\\
			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
			&\mbox{if } \alpha > 3/2.
		\end{cases}
\]
\end{lemma}

\begin{proof}
We will split the computation of $\Mu{\BallHyp{0,y}}$ into the cases $y^\prime > R_n - y$ and $y^\prime \le R_n - y$ where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ where $\xi_{\alpha,\nu} = \frac{4\alpha\nu}{\pi(2\alpha - 1)}$.

@@

For the finite box model $G_{\Pcal,n}(\alpha, \nu)$ we obtain a similar result for the average degree.

\begin{lemma}\label{lem:average_degree_P_n}
For all $p \in \Rcal_n$, such that $y > 2\log(\pi/2)$,
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - \phi_n(y)\right)
\]
where $\phi_n(y) \ge 0$ is given by
\[
	\phi_n(y) = \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)}e^{-(\alpha-\frac{1}{2})(R_n - y)}
	- \frac{(2\alpha - 1)\pi}{4\alpha}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} 
	e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(\alpha - \frac{1}{2})R_n - \frac{y}{2}}\right).
\]
On the other hand, if $y \le 2 \log(\pi/2)$ then
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - e^{-(\alpha - \frac{1}{2})R_n}\right).
\]
\end{lemma}

\begin{proof}
First note that since we have identified the boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ we can assume, without loss of generality, that $p = (0,y)$. We then have that
\[
	\BallPon{p} = \left\{p^\prime \in \Rcal_n \, : \, |x^\prime|_n \le e^{\frac{y + y^\prime}{2}}\right\},
\] 
whose boundaries given by the equations $x^\prime = \pm e^{\frac{y+y^\prime}{2}}$ intersect the left and right boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ at height
\[
	h(y) = R_n + 2 \log\left(\frac{\pi}{2}\right) - y.
\]
Therefore, if $y \le 2 \log(\pi/2)$ this intersection occurs above the height $R_n$ of the box $\Rcal_n$ while in the other case the full region of the box above $h(y)$ is connected to $p$.

@@

\subsection{Concentration argument for hyperbolic and finite box model}

@@

\begin{corollary}\label{cor:concentration_argument_other_models}
Let $\hat{\rho}_n(y,k)$ be any of the two distribution functions $\rho_{\HP,n}(y,k)$ and $\rho_{n}(y,k)$. Then, for any function $g_n(y)$ such that $g_n(y) = \bigO{n^{-1} k_n^{s} e^{\beta y}}\hat{\rho}_n(y,k_n)$, as $n \to \infty$, for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$,
\[
	\int_{\Rcal_n} g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y
	= (1+\smallO{1})\int_{\Kcal_C(k_n)} g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y,
\]
for some $C > 0$ large enough.
\end{corollary}

>>>>

Many of the computations in this paper will involve integrals whose integrand contains $\Prob{\Po(\mu_n(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and $\mu_n(y) \approx e^{y/2}$. For instance, the expected modified local clustering function $c_\infty(k)$ in $G_\Pcal$ equals
\[
	\frac{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \Delta_{\Pcal}(y) \alpha e^{-\alpha y} \dd y}
	{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \alpha e^{-\alpha y} \dd y},
\]
where $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$. 
\TM{ Maybe state where $\Delta$ is defined in the paper. }

@@

\subsection{Concentration \TM{ "of heights"? } argument for the infinite model}

Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xi_{\alpha,\nu}e^{\frac{y}{2}}$. 
For any sequence $\{k_n\}_{n \ge 1}$, possibly constant, define
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\end{equation}
\TM{ This is not a good definition. There are sequences that neither stay bounded nor tend to infinity, e.q.~odd values equal to 2 even values equal to $\log n$. How do you decide for a given $k_n$ which of the two cases you pick. To solve this we could either restrict to sequences that are either a) constant or b) tending to infinity; or we could make a cut-off at some specific, slow growing function such as $\log\log\log n$. }
and define further, for any $C > 0$
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(k_n) = \left\{p=(x,y) \in \R \times \R_+ : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
	\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},
\end{equation}
\TM{ taking max with 0 and min with $R$ do not seem necessary. }
Where we recall that $a \vee b = \max\{a,b\}$ and $a \wedge b = \min\{a,b\}$.

Note that if $k_n = \Omega(\log(n))$ then $e^{y/2} = \bigT{k_n}$ whenever $p=(x,y) \in \Kcal_C(k_n)$. The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{\Rcal_n} \rho(y,k_n) h(y) f_{\alpha,\nu}(x,y) \dd x \dd y
\]
\TM{ This part is supposed to be for the infinite model, yet you integrate over the finite box $\Rcal_n$. Is this really meant?
I would have expected we just integrate wrt.~$y$ and replace $f_{\alppha,\nu}(x,y)$ by $\alpha e^{-\alpha y}$, the exponential density.}
it is enough to consider integration \BL{over} $\Kcal_C(k_n)$ instead of $\Rcal_n$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \rightarrow  \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = \ell_n \pm C \kappa_n, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right),
\] 
\TM{ If $k$ is constant, then $\ell^- < 0$ and hence 
$a^-$ is undefined! Need to change the definitions. }

we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd x \dd y
	&= \begin{cases}
		\bigO{\log(n)^{k_n} n^{-\frac{C}{2}}} &\mbox{if } k_n = \bigT{1} \\
		\bigO{k_n^{-(1+C^2)/2}} &\mbox{else.}
	\end{cases}
\end{aligned}
\end{equation}
\TM{ As before, this is not a proper dichotomy. (unless we assume more about the sequences $k_n$, i.e.~either constant or tending to infinity }
as $n \to \infty$.

@@

\begin{proof}
Note that in this case
\[
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{\ell_n \log(\ell_n)} &\mbox{else.}
	\end{cases}
\]
Then, since $a_n^- < k_n$ and $a_n^+ > k_n$, $\rho_{y}(k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. 
\TM{ PLease give more details. }
Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho_{k_n}(y) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho_{k_n}(y) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{R_n} e^{\beta y} \rho_{k_n}(y) \alpha e^{-\alpha y}x \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho_{k_n}(y) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{R_n} \rho_{k_n}(y) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(y,a_n^+) \int_{a_n^+}^{R_n} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Note that if $k_n = \bigT{1}$ then, for large enough $n$, the first integral is zero. We conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho_{k_n}(y) \alpha e^{-\alpha y} \dd y
	&= \begin{cases}
		\bigO{1} \rho(a_n^+,k_n) &\mbox{if } k_n = \bigT{1},\\
		\bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) &\mbox{else.}
	\end{cases}	
\end{aligned}
\end{equation}

We shall now bound the terms $\rho(a_n^\pm,k_n)$, starting with $\rho(a_n^+,k_n)$.  Using that $k! > \sqrt{2\pi} k^{k + 1/2} e^{-k}$ \TM{ Reference! } we write
\begin{align*}
	\rho(a_n^+,k_n) &\BL{=} \frac{\Mu{\BallPo{a_n^+}}^{k_n}}{k_n!} e^{-\Mu{\BallPo{a_n^+}}} \\
	&\le (2\pi)^{-1/2} k_n^{-1/2} \left(\frac{\Mu{\BallPo{a_n^+}}}{k_n}\right)^{k_n} e^{-(\Mu{\BallPo{a_n^+}} - k_n)}\\
	&= (2\pi)^{-1/2} k_n^{-1/2} 
		e^{-k_n\left(\frac{\Mu{\BallPo{a_n^+}}}{k_n} - 1 - \log\left(\frac{\Mu{\BallPo{a_n^+}}}{k_n}\right)\right)}.
\end{align*}

Let us first consider the case $k_n \to \infty$,
\TM{ again, in general we cannot neatly split into the two cases $k\to\infty$ and $k=\Theta(1)$ } in which case $\kappa_n = \sqrt{\ell_n \log(\ell_n)}$. Since 
\[
	\frac{\Mu{\BallPo{a_n^+}}}{k_n} = \frac{\lambda_n^{\BL{+}}}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
Let use now consider the case $k_n = \bigT{1}$. Then $\kappa_n = \log(n)$,
\[
	\frac{\Mu{\BallPo{a_n^+}}}{k_n} = \frac{\lambda_n^+}{k_n} = 1 + \epsilon_n + C \frac{\log(n)}{k_n},
\] 
and hence
\begin{equation}\label{eq:concentration_lemma_bound_an+_constant}
	\rho(a_n^-,k_n) \le \bigO{\log(n)^{k_n} n^{-C}}.
\end{equation}

@@

\begin{remark}[Concentration argument]\label{rmk:concentration_argument}
\TM{ again, this is a loaded term. I would in the very least change to ``concentration of height'' everywhere }
Lemma~\ref{lem:concentration_argument} will prove very useful in the remainder of this paper since we often have to deal with integrands of the form $g_n(y) f_{\alpha,\nu}(x,y)$ where $g_n(y) = \bigO{n^{-1} k_n^{s}}e^{\beta y}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$. In this case the lemma tells us that for a suitable $C > 0$ we only need to integrate over $\Kcal_C(k_n)$. In order words, we may always assume for $g_n(y)$ (for a penalty of $\smallO{1}$) that $k_n - C \kappa_n \le \xi_{\alpha,\nu} e^{y/2} \le k_n + C \kappa_n$. We will refer to this as a \emph{concentration argument}, e.g. \emph{by a concentration argument}
\[
	\int_{\Rcal_n} g_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y 
	= (1+\smallO{1}) \int_{\Kcal_C(k_n)} g_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y.
\]
\PvdH{ @All:It might be a good idea to add one more example of how the concentration argument is used in the paper. I someone finds a good example we use later on, please add it here.}
\end{remark}

@@

\TM{  the term ``average degree" does not fit very well with what goes on in this section. We are considering the expected number of points in a ball, but not yet computing the average degree of the graph. }
Recall that under the coupling between the hyperbolic random graph and the finite box model $|x-x^\prime|_{\pi e^{R_n/2}} \le \Omega(r,r^\prime)$, while the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Omega(r, r^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for $y + y^\prime < R_n$. \TM{ Say what $K$ is and what the restrictions on $r, r'$ are. } This result enables us to determine the measure of a ball around a given point $p=(0,y)$ which will be fairly useful in our subsequent analysis.

\begin{lemma}\label{lem:average_degree_hyperbolic}
Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
\[
	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
\]
\TM{ maybe remind the reader just before of the rather counter intuitive definition of $\BallHyp{0,y}$, which 
lives in the box model not in the original hyperbolic disk. }
where
\[
	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 1 - 4\pi}{4\pi} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
		+ \left(\alpha - \frac{1}{2}\right)\pi e^{-(\alpha - \frac{1}{2})R_n - y/2},
\]
and
\[
	\phi_{\H,n}^{(2)}(y) = 
	+ \begin{cases}
			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
			&\mbox{if } 1/2 < \alpha < 3/2,\\
			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
			&\mbox{if } \alpha > 3/2.
		\end{cases}
\]
\TM{ We want to say in the lemma where $K$ comes from. (not everyone reads surrounding text, or reads linearly, in a paper as long as this one.) }
\end{lemma}

\begin{proof}
We will split the computation of $\Mu{\BallHyp{0,y}}$ into the cases $y^\prime > R_n - y$ and $y^\prime \le R_n - y$ \TM{ You have not said what $y^\prime$ is! } where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ where $\xi_{\alpha,\nu} = \frac{4\alpha\nu}{\pi(2\alpha - 1)}$.

@@

\TM{ same comment Re: average degree. }
For the finite box model $G_{\Pcal,n}(\alpha, \nu)$ we obtain a similar result for the average degree.

\begin{lemma}\label{lem:average_degree_P_n}
For all $p \in \Rcal_n$, such that $y > 2\log(\pi/2)$,
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - \phi_n(y)\right)
\]
\TM{ where did the curly B's go? }
where $\phi_n(y) \ge 0$ is given by
\[
	\phi_n(y) = \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)}e^{-(\alpha-\frac{1}{2})(R_n - y)}
	- \frac{(2\alpha - 1)\pi}{4\alpha}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} 
	e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(\alpha - \frac{1}{2})R_n - \frac{y}{2}}\right).
\]
On the other hand, if $y \le 2 \log(\pi/2)$ then
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - e^{-(\alpha - \frac{1}{2})R_n}\right).
\]
\end{lemma}

\begin{proof}
First note that since we have identified the boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ we can assume, without loss of generality, that $p = (0,y)$. We then have that
\[
	\BallPon{p} = \left\{p^\prime \in \Rcal_n \, : \, |x^\prime| \le e^{\frac{y + y^\prime}{2}}\right\},
\] 
\TM{ There a rogue subscript $n$ in $|x'|_n$ }
whose boundaries given by the equations $x^\prime = \pm e^{\frac{y+y^\prime}{2}}$ intersect the left and right boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ at height
\TM{ weird sentence }
\[
	h(y) = R_n + 2 \log\left(\frac{\pi}{2}\right) - y.
\]
Therefore, if $y \le 2 \log(\pi/2)$ this intersection occurs above the height $R_n$ of the box $\Rcal_n$ while in the other case the full region of the box above $h(y)$ is connected to $p$.

@@

\subsection{Concentration \TM{ OF HEIGHTS? }argument for hyperbolic and finite box model}

@@

\begin{corollary}\label{cor:concentration_argument_other_models}
Let $\hat{\rho}_n(y,k)$ be any of the two distribution functions $\rho_{\HP,n}(y,k)$ and $\rho_{n}(y,k)$. Then, for any function $g_n(y)$ such that $g_n(y) = \bigO{n^{-1} k_n^{s} e^{\beta y}}\hat{\rho}_n(y,k_n)$ \TM{ Where did the $n^{-1}$ come from? 
Does not appear in L3.4. }, as $n \to \infty$, for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$,
\[
	\int_{\Rcal_n} g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y
	= (1+\smallO{1})\int_{\Kcal_C(k_n)} g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y,
\]
for some $C > 0$ large enough.
\end{corollary}
